The Etiology of Cybersecurity

### 2020-07-16 reports/paper_0/main.tex
review fonts in images and add nonce example in paper_0

We now define them abstractly, inspired by the
ABF-theory  defined in \autocite{Santaca2016abf} and, in
Section~\ref{sec:theory} we detail them as engineering concepts.
\begin{enumerate}
	\item We consider Information only when the intention of exchanging 
		that Information is from a sender to 
		recipient is defined; and we call it \emph{Assertion}  
	\item Similarly, the portion of \emph{Beliefs} we consider for system
		engineering is the one that builds (input) or describes
		(output) the behavior of an agent (strategy rules
		\footnote{``The logical structure of information is one of the
		most basic and one of the most basic and one of the simplest
		thing in the wide and wonderful world of logical analysis. This
		point can be put in a deeper perspective. A distinction
		[\ldots] ought to be made [--] between two kinds of rules (or
		principles) in any strategic activity like knowledge seeking.
		On the one hand you have the rules that define the game, e.g.
		how chessmen are moved on a board. The can be called
		\emph{definitory} rules.  They must be distinguished from rules
		[\ldots] that deal with what is better and what is worse in the
		game in question.  Definitory rules do not say anything about
		this subject. Rules which do can be called \emph{strategic
		rules}'' -- Hintikka in \autocite{Hintikka1993Information}}),
		and
	\item We consider a set of axiomatic \emph{Facts} (definitory
		rules\footnote{Facts are definitory rule as they don't
		define how a real system is finally implemented but
		how it should be, trough a series of requirements. Those requirements 
		may not hold, through insecurity, in the implemented system.})
		instead of considering the more general epistemic definition of
		Knowledge. Specifically, Facts describes:
		\begin{itemize}
			\item The functional architecture of each agent
			\item The physical/structural (HW/SW) architecture of each
				subsystem of agents and agent within a
				subsystem
		\end{itemize}
\end{enumerate}
We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.
The notation will be introduced in the next section but, intuitively,
Figure~\ref{fig:system-agent} shows a system composed by two agents. The system
has to ports denoted by ``I'' (with incoming assertions) and ``O'' (with outgoing assertions).
The bottom part of Figure~\ref{fig:system-agent} details the structure of a single
agent which is not decomposed into other (sub-)agents. In this case, ports translates
incoming assertions into beliefs that feeds the behavior block, and outgoing beliefs
from the behavior block are then translated back into (outgoing) assertions.

@@

Hence, even if considering a different number of regions than the three
$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
the number of theoretical agents, the application of RCC downscales that number
of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
3.6 times more (different) types of agents than RCC3, but using RCC8 would
allow us to consider 3.5 times more different agents.
In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior. We note 
that, the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system.
\fixnote{mv}{In Fig. 5 e 6, i simboli A, B e F hanno un font diverso dal solito}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{quantitative.pdf}
	\caption{Cybersecurity risk for a single agent}
	\label{fig:quantitative}
\end{figure}

@@

\subsubsection{$\abf$ Security Enumeration (SE)} The following security
requirement for a CPS specification can be summarized:
\begin{enumerate}
	\item[SE-$1$] Proper interaction between correctly-behaving agents is
		defined as $\eq{\assertionRegion_a}{\beliefRegion_a}$ for an
		agent $a$ while can be detailed as follows when multiple agents
		are considered.
	\begin{enumerate}
		\item[SE-$1.1$] The equality relation
			$\eq{\assertionRegion_{s\rightarrow
			a}}{\beliefRegion_a}$ describes the intended secure
			behavior as: the beliefs generated by the behavior of
			the functional architecture shall be complete w.r.t.
			the specified inputs of the agent. Therefore, \emph{the
			assertions received by an agent or a system shall be
			compliant with the expected inputs of the functional
			architecture}. For example, the inputs of the user of a
			SW must be sanitized to exclude deviations w.r.t. the
			expected inputs of the functions implemented in the SW.
			Another example is the type checking between allowed
			inputs and expected inputs.
		\item[SE-$1.2$] Similarly, the equality relation
			$\eq{\assertionRegion_{a\rightarrow
			r}}{\beliefRegion_a}$ defines that the outputs of an
			agent $a$ shall be the outputs of the functional
			architecture.
	\end{enumerate}
\item[SE-$2$]\fixnote{mv}{Are agents missing in the notation A,B,F for SE-2 and SE-3?}
	Sufficient Control Flow Management  is defined as
		$\eq{\assertionRegion}{\factRegion}$.
	\item[SE-$3$] Correct Calculation  is defined as
		$\eq{\beliefRegion}{\factRegion}$.
\end{enumerate}

@@

We are now in the position to define what a secure system is (with respect to
the $\abf$-theory) and, based on that definition, what the security risk is and
how to quantify it in a risk matrix (where a risk matrix is usually defined as
a function of the likelihood and the impact of all potential attacks).
\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
	each agent composing the system.
\end{definition}
The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
refers both to positive and negative consequences of uncertainty.
Accordingly, we consider risk as follows. 
\begin{definition}{\bf Risk --}
A risk is the whole space of potential designs of a specification with respect to the
	$\abf$-theory.
\end{definition}
The definition of Risk leads to the risk matrix in
Figure~\ref{fig:quantitative}, defined as follows.
\begin{definition}{\bf Risk Matrix --}
	The risk matrix, as summarized in Table~\ref{tab:5com}, is a function of the three relations
	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where the maximum risk is defined by the $DR$ relation between the
	three groups of Regions, and the minimum risk by the $EQ$ relation over
	the same Regions. In between the two extremes, the granularity of
	possible intermediate configuration is defined by the calculus used
	(RCC5 in our case).
\end{definition}
While a risk matrix is often defined as a function of 
likelihood and impact of attacks, we suggest
that security weaknesses are all equally likely to be exploited
if there's a connection between the weakness and the asset that
an attacker wants to impact. Therefore, a risk matrix should
capture the number of insecure configurations of a system, rather
than predicating over likelihoods of weaknesses exploitation.

\section{Prediction of Security Weaknesses}\label{sec:theory}
Several standards mandates a secure-by-design approach in which cybersecurity
shall be considered at the very early stages of the design process.  For
example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and is
		the ``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
However, standards do not describe in detail how to perform a security risk 
assessment and only vaguely define the overall objective. The overall
objective of this first step of the cybersecurity engineering process
is to provide an understanding of the potential security risks.
Roughly speaking, a security risk assessment (e.g. CORAS\autocite{Lund2010model}) 
methodology starts from the (i) identification of assets, (ii) determines
the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
threat diagrams in the CORAS
approach, (iii) calculate the risk as a function
of impact and likelihood of threat scenarios, and (iv) finally provides
treatments (mandating a partial re-design) and residual risks.
All the methodologies and tools we reviewed (e.g. Threatmodeler \autocite{Threatmodeler},
CORAS \autocite{Lund2010model}, SECRAM \autocite{De2015role})
relies on the expertise of the person who performs the risk assessment for
the identification of threats and for the quantitative estimation of risks.
As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
we don't have a metric for security'' while standards requires a security risk assessment
which, in turn, mandates a metric to evaluate the security risk.

@@

\paragraph{Input and Output Ports}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{IOports.pdf}
	\caption{Input and Output Port of an agent}
	\label{fig:ioports}
\end{figure}
Since the $\abf$-theory is a theory of agents, we could consider ports as
agents that allows the exchange of information between a channel and another
agent.  However, we considered a port as a special type of agent that to avoid
an \emph{infinite regress} in which a port needs a port to transfer information
between the outside of the port to the inside of itself.

@@

\begin{definition}{\bf Secure Input Port --}\label{def:secport}
	A secure input port ($s\secureport a$) always allow information as
	incoming assertions to flow from a sender $s$ to the behavior of the
	recipient $a$ (agent).  
	%If $p_I$ is a secure input port,
	%$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$ has the
	%only input port $p_I$, $\world'\models\beliefRegion_a$ and
	%$\varphi\in\beliefRegion_a$ for any $\world\modalrelation\world'$.
\end{definition}

@@

\begin{proof}
An input port is, in the $\abf$-theory, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different type
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Lets
	define a function weight $|X|$ such as, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every open region can be written as union of elements of that base. 
	We distinguish between Regions that contain Information and Regions that
	don't by writing the latter as $\emptyset$.
	\begin{enumerate}
		\item if $\eq{\assertionRegion}{\beliefRegion})$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication)
		\item if $\dr{\assertionRegion}{\beliefRegion})$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former and \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			which we call \emph{replace} (i.e. drop and insert)
		\item if $\pp{\assertionRegion}{\beliefRegion})$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{injection}
		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{selective drop}
		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation}
	\end{enumerate}
\end{proof}

@@

We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore,
we consider channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a axiomatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}.

@@

\paragraph{Security Weaknesses} 
It is important to note that all the results of the application of the
$\abf$-theory to channels (the analysis of the relation
$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
analysis of a pair of an input and output port (i.e.
$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
$\Rcc{\assertionRegion_2}{\beliefRegion_3}$).  The same result can be obtained
by analyzing the relation between the outputs of a functional block and the
inputs of another functional block ($\Rcc{\beliefRegion_o}{\beliefRegion_i}$), 
where functional blocks are constituents of
the functional architecture as described afterwards.  As depicted in
Figure~\ref{fig:atom}, so far, we have considered Information generated by a
Port $P_I$ and then sent through a channel $C$ to another (recipient) port
$P_O$. In this scenario where Ports and then Channels are atomic (otherwise
raising infinite regress), we can only  consider the relations between Ports
and Channel; considering both input-port to channel and channel to output-port.
In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
Ports. We now introduce the concept of functional architecture and its
decomposition into functional blocks.  Io order to define a functional block
without encountering a infinitely recursive definition, we must reach the same
conclusions as for the Channel. Therefore, describing the information as
flowing over a channel or in a functional block is purely syntactic sugar.

We can summarize these results by saying that the relations between Assertions
and Beliefs, Output Assertions of an Agent and Input Assertions of another
Agent, or Output Beliefs of a Functional Block and Input Beliefs of another
Block can \emph{only} be affected by the following weaknesses: replace, drop,
injection, insertion, selective drop, and selective drop + insertion.  We can,
then, deduce the following security properties to mitigate security weaknesses
of a port or a channel (between ports, functional blocks, or both).
\begin{itemize}
	\item \emph{Order-preserving} -- it shall be known if Information is \emph{replaced}
	\item \emph{Availability} -- it shall be known if Information is \emph{dropped} or \emph{selectively dropped}
	\item \emph{Integrity} -- it shall be known if Information is \emph{injected}
	\item \emph{Authentication} -- it shall be known if Information is \emph{inserted}
\end{itemize}
%We can also generalize the weaknesses related to the data flow (i.e. the flow
%of Information as Assertions or Beliefs) as the relation between output-input
%Beliefs between functional blocks, output-input Assertions of a channel,
%inputs-outputs of a port, as follows.
%\begin{itemize}
%	\item \emph{equal}, expected nominal behavior
%	\item \emph{discrete-from}, drops all the inputs and inserts new malicious data
%	\item \emph{proper part}, selectively drops inputs
%	\item \emph{proper part inverse}, forwards all the inputs but crafts and
%	      inserts new malicious data
%	\item \emph{partial overlap}, selectively drops inputs and inserts new data.
%\end{itemize}

@@

\paragraph{Assertions and Facts}
The whole reasoning on the relation between Beliefs and Facts can be duplicated
for the relation between Assertions and Facts; we cannot appreciate the
difference at this level of abstraction.  If the functional architecture would
be extended to capture the semantics (i.e. the logic) of the communication and
security protocols, with the relation between Assertions and Facts we would
compare protocol logics with the requirements. We consider the verification of
the functional architecture and protocol logic in subsequent steps of the
engineering process (as we describe in Section~\ref{sec:discussion}).
Therefore, we can summarize our results by categorizing the weaknesses
predicted by our theory into: data-flow-related and functionality-related
weaknesses; as in Table~\ref{tab:weaknesses}.
\begin{table*}[t]
\centering
	\begin{tabular}{c C{.35\textwidth} C{.5\textwidth}} 
		\textbf{RCC5} & \textbf{Quantity: Data Flow} & \textbf{Quality: Requirements Adherence}\\
		\hline 
		EQ & Expected/Nominal & Expected/Nominal\\[.1cm]
	DR & Drops all the inputs and inserts new malicious data & The component never performs/carries the expected behavior/information\\[.1cm]
	PP & Selectively drops inputs & Part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
	PPi& Forwards all the inputs but crafts and inserts new malicious data & The components correctly performs/carries the expected behavior/information when the correct inputs are provided but is subject to input injections \\[.1cm]
	PO & Selectively drops inputs and inserts new data & Byzantine behavior - occasionally outputs the expected output given the correct inputs. Not all the inputs are handled properly, nor all the expected outputs are always generated when correct inputs are given
\end{tabular}
\caption{Weaknesses Categorized~\label{tab:weaknesses}}
\end{table*}

\subsection{Security and Insecurity of a System}
\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
	A system security design is given by a precise system 
	specification over the physical and functional architectures, that
	uniquely, i.e.  tested against the range of design possible in the
	$\abftheory$-theory, defines the design built on top of those
	requirements.
\end{hypothesis}

>>>>

We now define them abstractly, inspired by the
$\abf$-framework  defined in \autocite{Santaca2016abf} and, in
Section~\ref{sec:theory} we detail them as engineering concepts.
\begin{enumerate}
	\item We consider Information only when the intention of exchanging 
		that Information is from a sender to 
		recipient is defined; and we call it \emph{Assertion}  
	\item Similarly, the portion of \emph{Beliefs} we consider for system
		engineering is the one that builds (input) or describes
		(output) the behavior of an agent (strategy rules
		\footnote{``The logical structure of information is one of the
		most basic and one of the most basic and one of the simplest
		thing in the wide and wonderful world of logical analysis. This
		point can be put in a deeper perspective. A distinction
		[\ldots] ought to be made [--] between two kinds of rules (or
		principles) in any strategic activity like knowledge seeking.
		On the one hand you have the rules that define the game, e.g.
		how chessmen are moved on a board. The can be called
		\emph{definitory} rules.  They must be distinguished from rules
		[\ldots] that deal with what is better and what is worse in the
		game in question.  Definitory rules do not say anything about
		this subject. Rules which do can be called \emph{strategic
		rules}'' -- Hintikka in \autocite{Hintikka1993Information}}),
		and
	\item We consider a set of axiomatic \emph{Facts} (definitory
		rules\footnote{Facts are definitory rule as they don't
		define how a real system is finally implemented but
		how it should be, trough a series of requirements. Those requirements 
		may not hold, through insecurity, in the implemented system.})
		instead of considering the more general epistemic definition of
		Knowledge. Specifically, Facts describes:
		\begin{itemize}
			\item The functional architecture of each agent
			\item The physical/structural (HW/SW) architecture of each
				subsystem of agents and agent within a
				subsystem
		\end{itemize}
\end{enumerate}
We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.
The notation will be introduced in the next section but, intuitively,
Figure~\ref{fig:system-agent} shows a system composed by two agents. The system
has to ports denoted by ``I'' (with incoming assertions) and ``O'' (with outgoing assertions).
The bottom part of Figure~\ref{fig:system-agent} details the structure of a single
agent which is not decomposed into other (sub-)agents. In this case, ports translates
incoming assertions into beliefs that feeds the behavior block, and outgoing beliefs
from the behavior block are then translated back into (outgoing) assertions.

@@

Hence, even if considering a different number of regions than the three
$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
the number of theoretical agents, the application of RCC downscales that number
of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
3.6 times more (different) types of agents than RCC3, but using RCC8 would
allow us to consider 3.5 times more different agents.
In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior. We note 
that, the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system.
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{quantitative.pdf}
	\caption{Cybersecurity risk for a single agent}
	\label{fig:quantitative}
\end{figure}

@@

\subsubsection{$\abf$ Security Enumeration (SE)} The following security
requirement for a CPS specification can be summarized:
\begin{enumerate}
	\item[SE-$1$] Proper interaction between correctly-behaving agents is
		defined as $\eq{\assertionRegion_a}{\beliefRegion_a}$ for an
		agent $a$ while can be detailed as follows when multiple agents
		are considered.
	\begin{enumerate}
		\item[SE-$1.1$] The equality relation
			$\eq{\assertionRegion_{s\rightarrow
			a}}{\beliefRegion_a}$ describes the intended secure
			behavior as: the beliefs generated by the behavior of
			the functional architecture shall be complete w.r.t.
			the specified inputs of the agent. Therefore, \emph{the
			assertions received by an agent or a system shall be
			compliant with the expected inputs of the functional
			architecture}. For example, the inputs of the user of a
			SW must be sanitized to exclude deviations w.r.t. the
			expected inputs of the functions implemented in the SW.
			Another example is the type checking between allowed
			inputs and expected inputs.
		\item[SE-$1.2$] Similarly, the equality relation
			$\eq{\assertionRegion_{a\rightarrow
			r}}{\beliefRegion_a}$ defines that the outputs of an
			agent $a$ shall be the outputs of the functional
			architecture.
	\end{enumerate}
\item[SE-$2$] The proper adherence of the data transmitted between agents with respect to
	requirements, is defined as $\eq{\assertionRegion}{\factRegion}$. 
	\item[SE-$3$] The proper adherence of the behavior with respect to
		requirements is defined as $\eq{\beliefRegion}{\factRegion}$.
\end{enumerate}

@@

We are now in the position to define what a secure system is (with respect to
the $\abf$-framework) and, based on that definition, what the security risk is and
how to quantify it in a risk matrix.
\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
	each agent composing the system.
\end{definition}
The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
refers both to positive and negative consequences of uncertainty.
Accordingly, we consider risk as follows. 
\begin{definition}{\bf Risk --}
A risk is the whole space of potential designs of a specification with respect to the
	$\abf$-framework.
\end{definition}
The definition of Risk leads to the risk matrix in
Figure~\ref{fig:quantitative}, defined as follows.
\begin{definition}{\bf Risk Matrix --}
	The risk matrix, as summarized in Table~\ref{tab:5com}, is a function of the three relations
	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where the maximum risk is defined by the $DR$ relation between the
	three groups of Regions, and the minimum risk by the $EQ$ relation over
	the same Regions. In between the two extremes, the granularity of
	possible intermediate configuration is defined by the calculus used
	(RCC5 in our case).
\end{definition}
While a risk matrix is often defined as a function of the 
likelihood and impact of attacks (based on quantitative ad-hoc estimation
of how likely it is that an attacker will exploit one or more vulnerabilities
and what is the magnitude of the incidents produced by this exploitation), we suggest
that security weaknesses are all equally likely to be exploited
if there's a connection between the weakness and the asset that
an attacker wants to impact. Therefore, a risk matrix should
capture the number of insecure configurations of a system, rather
than predicating over likelihoods of weaknesses exploitation.

\section{Prediction of Security Weaknesses}\label{sec:theory}
Several standards mandates a secure-by-design approach in which cybersecurity
shall be considered at the very early stages of the design process.  For
example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and is
		the ``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
However, standards do not describe in detail how to perform a security risk 
assessment and only vaguely define the overall objective, which
can be summarized as to provide an understanding of the potential security risks.
Roughly speaking, a security risk assessment (e.g. CORAS\autocite{Lund2010model}) 
methodology starts from the (i) identification of assets, (ii) determines
the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
threat diagrams in the CORAS
approach, (iii) calculate the risk as a function
of impact and likelihood of threat scenarios, and (iv) finally provides
treatments (mandating a partial re-design) and residual risks.
All the methodologies and tools we reviewed (e.g. Threatmodeler \autocite{Threatmodeler},
CORAS \autocite{Lund2010model}, SECRAM \autocite{De2015role})
relies on the expertise of the person who performs the risk assessment for
the identification of threats and for the quantitative estimation of risks.
As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
we don't have a metric for security'' while standards requires a security risk assessment
which, in turn, mandates a metric to evaluate the security risk.

@@

\paragraph{Input and Output Ports}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{IOports.pdf}
	\caption{Input and Output Port of an agent}
	\label{fig:ioports}
\end{figure}
Since the $\abf$-framework is a theory of agents, we could consider ports as
agents that allows the exchange of information between a channel and another
agent.  However, we considered a port as a special type of agent that to avoid
an \emph{infinite regress} in which a port needs a port to transfer information
between the outside of the port to the inside of itself.

@@

\begin{definition}{\bf Secure Input Port --}\label{def:secport}
	A secure input port always allow information as
	incoming assertions to flow from a sender $s$ to the behavior of the
	recipient $a$ (agent).  
	%If $p_I$ is a secure input port,
	%$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$ has the
	%only input port $p_I$, $\world'\models\beliefRegion_a$ and
	%$\varphi\in\beliefRegion_a$ for any $\world\modalrelation\world'$.
\end{definition}

@@

\begin{proof}
An input port is, in the $\abf$-framework, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different type
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Lets
	define a function weight $|X|$ such as, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every open region can be written as union of elements of that base. 
	We distinguish between Regions that contain Information and Regions that
	don't by writing the latter as $\emptyset$.
	\begin{enumerate}
		\item if $\eq{\assertionRegion}{\beliefRegion})$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication)
		\item if $\dr{\assertionRegion}{\beliefRegion})$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former and \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			which we call \emph{replace} (i.e. drop and insert)
		\item if $\pp{\assertionRegion}{\beliefRegion})$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{injection}
		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{selective drop}
		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation}
	\end{enumerate}
\end{proof}

@@

We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abf$-framework is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore,
we consider channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a axiomatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}.

@@

\paragraph{Security Weaknesses -- The RIDI-Hypothesis} 
It is important to note that all the results of the application of the
$\abf$-framework to channels (the analysis of the relation
$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
analysis of a pair of an input and output port (i.e.
$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
$\Rcc{\assertionRegion_2}{\beliefRegion_3}$).  The same result can be obtained
by analyzing the relation between the outputs of a functional block and the
inputs of another functional block ($\Rcc{\beliefRegion_o}{\beliefRegion_i}$), 
where functional blocks are constituents of
the functional architecture as described afterwards.  As depicted in
Figure~\ref{fig:atom}, so far we have considered Information generated by a
Port $P_I$ and then sent through a channel $C$ to another (recipient) port
$P_O$. In this scenario where Ports and then Channels are atomic (otherwise
raising infinite regress), we can only  consider the relations between Ports
and Channel; considering both input-port to channel and channel to output-port.
In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
Ports.  Io order to define a functional block
without encountering a infinitely recursive definition, we must reach the same
conclusions as for the Channel. Therefore, describing the information as
flowing over a channel or in a functional block is purely syntactic sugar.

We can summarize these results by saying that the relations between Assertions
and Beliefs, Output Assertions of an Agent and Input Assertions of another
Agent, or Output Beliefs of a Functional Block and Input Beliefs of another
Block can \emph{only} be affected by the following weaknesses: replace, drop,
injection, insertion, selective drop, and selective drop + insertion. 
We call this the RIDI-Hypothesis, being the
four main categories of weaknesse: Replace, Insertion, Drop, Injection. We can,
then, deduce the following security properties to mitigate security weaknesses
of a port or a channel (between ports, functional blocks, or both).
\begin{itemize}
	\item \emph{Order-preserving} -- it shall be known if Information is \emph{replaced}
	\item \emph{Availability} -- it shall be known if Information is \emph{dropped} or \emph{selectively dropped}
	\item \emph{Integrity} -- it shall be known if Information is \emph{injected}
	\item \emph{Authentication} -- it shall be known if Information is \emph{inserted}
\end{itemize}
%We can also generalize the weaknesses related to the data flow (i.e. the flow
%of Information as Assertions or Beliefs) as the relation between output-input
%Beliefs between functional blocks, output-input Assertions of a channel,
%inputs-outputs of a port, as follows.
%\begin{itemize}
%	\item \emph{equal}, expected nominal behavior
%	\item \emph{discrete-from}, drops all the inputs and inserts new malicious data
%	\item \emph{proper part}, selectively drops inputs
%	\item \emph{proper part inverse}, forwards all the inputs but crafts and
%	      inserts new malicious data
%	\item \emph{partial overlap}, selectively drops inputs and inserts new data.
%\end{itemize}

@@

\paragraph{Assertions and Facts}
The whole reasoning on the relation between Beliefs and Facts can be duplicated
for the relation between Assertions and Facts; we cannot appreciate the
difference at this level of abstraction.  If the functional architecture would
be extended to capture the semantics (i.e. the logic) of the communication and
security protocols, with the relation between Assertions and Facts we would
compare protocol logics with the requirements. We won't consider the verification of
the functional architecture and protocol logic in this paper since we
focus on the architecture specification step of the engineering process
without going into the design of the behavior of agents. We can give an example
but research is needed to apply our theory
to the behavior of agents.

If we consider a functional block that generates nonces (i.e. number used
only once), the semantics of this block must define a procedure such that 
a freshness requirement holds. By the RIDI-Hypothesis we have that
the following weaknesses may occur:
\begin{itemize}
	\item Replace: the output-nonces of the block are replaced with
		non-fresh numbers
	\item Insert: the output-nonces are concatenated with other numbers,
		resulting in non-fresh nonces. For example, the nonces 110 and
		111 are modified by inserting a 1 and a 0 at the beginning and
		end respectively, resulting in the two identical (and the
		non-fresh) nonces 1110 and 1110
	\item Delete: part of output-nonce is dropped. For example, the last bit of 1101 and 1100 is dropped, resulting in the two equal and non-fresh nonces: 110 and 110.
	\item Inject: substitution at bit level results in non-fresh nonces. For example, 110 becomes 111 becomes equal by flipping either the last 0 of the former nonce or the last 1 of the latter.
\end{itemize}

We summarize our results by categorizing the weaknesses
predicted by our theory into: data-flow-related and functionality-related
weaknesses; as in Table~\ref{tab:weaknesses}.
\begin{table*}[t]
\centering
	\begin{tabular}{c C{.35\textwidth} C{.5\textwidth}} 
		\textbf{RCC5} & \textbf{Quantity: Data Flow} & \textbf{Quality: Requirements Adherence}\\
		\hline 
		EQ & Expected/Nominal & Expected/Nominal\\[.1cm]
	DR & Drops all the inputs and inserts new malicious data & The component never performs/carries the expected behavior/information\\[.1cm]
	PP & Selectively drops inputs & Part of the expected outputs are not generated in response to the correct inputs \\[.1cm]
	PPi& Forwards all the inputs but crafts and inserts new malicious data & The components correctly performs/carries the expected behavior/information when the correct inputs are provided but is subject to input injections \\[.1cm]
	PO & Selectively drops inputs and inserts new data & Byzantine behavior - occasionally outputs the expected output given the correct inputs. Not all the inputs are handled properly, nor all the expected outputs are always generated when correct inputs are given
\end{tabular}
\caption{Weaknesses Categorized~\label{tab:weaknesses}}
\end{table*}
